Supplementary Material for `The response of trophic interaction networks to multiple stressors in a marine latitudinal gradient of the Southern Hemisphere'

### 2024-04-18 manuscript.Rmd
Ms reorganization almost to submit

# The response of trophic interaction networks to multiple stressors in a marine latitudinal gradient of the Southern Hemisphere
## Running Title: Marine stressors in the Southern Hemisphere

@@

Ecological networks offer valuable insights into community structure, key species identification, and ecosystem management for biodiversity conservation. Understanding how these networks react to environmental and anthropogenic stressors, especially along geographical gradients, is of increasing interest.
This review presents a pioneering analysis of stressor responses in marine food webs from the southwest Atlantic to the Antarctic (45 - 78ºS), encompassing areas such as San Jorge Gulf, Beagle Channel, Burdwood Bank, Scotia Sea, Potter Cove, and the Weddell Sea in Antarctica. Our objectives are to: 1) describe the structure of marine food webs along this axis using a network approach; 2) identify predominant environmental and anthropogenic stressors affecting each ecosystem; and 3) summarize observed food web changes and hypothesize on stressor impacts.
Our collaborative team, comprising regional experts and global authorities on high-latitude marine food webs and stressor effects, ensures a comprehensive and credible literature review. We assessed the effects of stressors primarily at the species level, with notable exceptions like fisheries in San Jorge Gulf.
Hypotheses for each study area were formulated considering: a) stressors; b) impacted parameters; c) node-level species properties; and d) network-level food web properties. Global warming emerges as the most common stressor across the gradient, except in the Beagle Channel and Burdwood Bank, where alien species introduction and fisheries are more influential, respectively. We offer specific hypotheses on how warming may affect food webs.
Our findings highlight the benefits of a network approach in understanding and predicting stressor effects in Southern Hemisphere marine ecosystems. This approach provides a holistic understanding of ecological networks, enhances our ability to identify key species and interactions, and offers insights for ecosystem management and conservation in the face of various stressors.

Keywords: anthropogenic stressors, environmental stressors, food webs, latitudinal gradient, Southern Hemisphere

@@

The application of a network perspective has emerged as a powerful tool to tackle the complexity of species interactions, facilitating a better understanding of the structure and functioning of ecosystems [@Belgrano2005; @Thompson2012]. Trophic networks (or food webs) allow identifying properties and key species that may be crucial for ecosystem stability, and hence important for ecosystem management and biodiversity conservation [@Thompson2012].
There is a growing interest in understanding how ecological networks respond to environmental and anthropogenic stressors along geographical gradients [@Bauer2022; @Cirtwill2015]. Yet, only a few studies have described variation in food web structure along latitudinal gradients in marine ecosystems. The few that have come from the Global North [@Wood2015; @Kortsch2019; @Pecuchet2022], whereas no studies, nor meta-analyses, on geographical variation in marine food webs exist for the Global South (Southern Hemisphere).

Here we review for the first time the state-of-the-art knowledge on stressor response of marine food webs along the southwest Atlantic to Antarctic gradient (45 - 78ºS, Figure 1). We focus on proven and expected changes in food webs driven by stressors in selected areas along this large-scale latitudinal gradient. We recruited food web experts from different marine systems of Argentina and the world (see co-authors’ list). Throughout the year 2023 we maintained regular discussion meetings, typically held every two or three weeks. The aim of this review is threefold: 1) describe the complexity and structure of marine food webs along the southwest Atlantic to Antarctic axis from a network perspective; 2) identify the ongoing environmental and anthropogenic stressors for each marine ecosystem containing the food webs; and 3) review proven food web changes and elaborate hypotheses on how the identified stressors might affect food web features (e.g. energy flow, stability), combining information on node- and network-level properties. Finally, we suggest which additional data and analyses are needed to gain insights into the stressors’ effects on food web properties in the southwest Atlantic to Antarctic region.

## 2. The structure of marine food webs in the Southwest Atlantic - Antarctic region

Together, the southwest and the Atlantic sector of the Southern Ocean comprise one of the most biologically productive regions of the world's oceans [@Acha2004;  @Latorre2023]. The referred region extends from San Jorge Gulf (45ºS) in the Patagonian shelf to the Weddell Sea (78ºS) in the Southern Ocean, and covers a well-connected oceanic latitudinal gradient [@Matano2010; @Guihou2020].

Throughout this latitudinal gradient, many investigations have been carried out addressing the trophic ecology of specific species and prey-predator relationships [@Vinuesa2007; @Pasotti2015;  @Riccialdelli2020; @Saunders2019], but few studies have considered the complexity of the ecosystem in terms of a high resolution of species and their prey-predator interactions (but see @Jacob2011, @Marina2018, @Funes2022, @Lopez-Lopez2022, @Rodriguez2022, @Marinainrev). Neglecting this complexity might lead to a misunderstanding of the structure and functioning of the ecosystems, and ultimately reduce the ability to predict ecosystem responses to perturbations [@Montoya2009].

In this review, we consider marine areas in the Southern hemisphere for which highly-resolved empirical food webs, in terms of species diversity and trophic interactions, have been previously described. These areas include: (1) San Jorge Gulf (45 - 47ºS, 65 - 68ºW), (2) Beagle Channel (~54ºS, 68ºW), (3) Burdwood Bank (~54ºS, 59ºW), (4) Scotia Sea (48 - 58ºS, 50ºW), (5) Potter Cove (62ºS, 58ºW, Antarctica), and (6) Weddell Sea (74 - 78ºS, 30ºW, Antarctica) (Figure 1). The geographic locations of these marine biomes span from temperate to Antarctic, and are exposed to both environmental (sea warming, glacial retreat) and anthropogenic (fishery, pollution) stressors.

@@

San Jorge Gulf is the northernmost study area considered in this review. It is a partially enclosed basin spanning approximately 34,000 km2 and ~100 m of maximum depths, located from Cabo dos Bahias (44°55'S) to Cabo tres Puntas (47°06'S) (Figure 1). The shallower northern and southern ends of the Gulf (with depths ~40 m) present two prominent frontal systems. These systems are the areas of highest productivity [@Glembocki2015]. The Gulf’s productivity supports large invertebrate and vertebrate fisheries [@Gongora2012], as well as marine mammal and seabird populations [@Yorio2009].
The San Jorge Gulf food web comprises 165 nodes and 1015 trophic interactions with a connectance of 0.04. The percentage of top predators is 16%, 78% of intermediate nodes and 6% of basal nodes; 60% of predators are omnivorous (Table 1). The most connected nodes are: the Argentine red shrimp Pleoticus muelleri, the squat lobster Gimothea gregaria, squids (Illex argentinus as dominant species) and Amphipoda. Notably, these nodes present mid-trophic positions in the food web (3.0, 2.5, 3.6 and 2, respectively) [@Funes2022].

At a southern latitude, the Beagle Channel (54°52’S, 68°8’W) is an interoceanic passage located at the southernmost tip of South America, spanning 240 km in length and 5 km in width (~1200 km2), and a range depth of 0 - 140 m (Figure 1). It features complex coastlines, varying bathymetry, a prevailing west-to-east circulation pattern, and a significant longitudinal gradient of glacial freshwater discharge [@Schloss2023]. @Bruno2023 suggests that locally produced suspended particulate organic matter (mostly composed by phytoplankton) and organic matter accumulated in the sediments (macroalgae-originated detritus) are the primary food sources for the marine ecosystem, as opposed to allochthonous materials.
Beagle Channel food web includes 145 nodes and 1115 trophic interactions with a connectance of 0.05 (Table 1). The food web is suggested to have a wasp-waist structure, where the following species play a crucial role in the dynamics of the ecosystem: Fueguian sprat Sprattus fuegensis, longtail southern cod Patagonotothen ramsayi, black southern cod P. tessellata, frogmouth Cottoperca trigloides, and squat lobster Grimothea gregaria [@Riccialdelli2020]. Moreover, the squat lobster has been identified as a species responsible for linking modules and connecting the entire food web [@Rodriguez2022].

The so-called Burdwood Bank ecosystem comprises Marine Protected Areas Namuncurá - Burdwood Bank I and II, meaning the shallow submarine plateau named Burdwood Bank with a 200 m isobath boundary, and a deep slope that reaches 4000 m in depth, respectively [@APN2022] (Figure 1). Physical features in the plateau are fairly stable, with salinity averaging 34 all year round and temperature ranging between 4 and 8ºC [@Acha2004]. The plateau is surrounded by steep flanks of up to 4000 m depth, protected by the Marine Protected Area Namuncurá - Burdwood Bank II (32,000 km2; ~55º - 56ºS, ~58º - 62ºW). Intense upwelling and mixing occur in relation to  the slope, entraining deep nutrient-rich waters into the photic layer [@Matano2019], and resulting in a fairly homogeneous water column both spatially and temporally [@Matano2019].
The Burdwood Bank food web comprises 379 nodes and 1788 interactions, with a connectance of 0.01, and an asymmetric degree distribution (i.e. most of the species have a relatively low number of interactions and few species concentrate most of them). Almost half of the consumers are omnivores (0.48), and the network displays a small-world pattern [@Marinainrev] (Table 1).

The Scotia Sea is a deep-sea basin, delimited by the Drake Passage to the West and by the island complex of the Scotia Arc to the North, East, and South, with an approximate extension of 1.5 x 106 km2 and a depth range of 0 - 3000 m [@Murphy2006] (Figure 1). Its oceanography is dominated by the Antarctic Circumpolar Current, which is spatially structured by frontal systems [@Whitworth1980]. The South Antarctic Circumpolar Current Front subdivides the Scotia Sea into two biogeographic regions: the Northern Scotia Sea is characterized by higher and more variable temperatures, and the Southern Scotia Sea by lower and more stable temperatures and influenced by seasonal sea ice [@Raymond2011].
The analysis of the Northern and Southern Scotia Sea food webs shows that the former is relatively more complex than the latter: with higher species richness (218 vs 192) and interactions (10008 vs 7241) and a higher connection overall (0.21 vs 0.20). As expected from a more complex network, the path length is shorter in the Northern Scotia Sea food web. In the same sense, the Southern Scotia Sea network displays a greater proportion of omnivores and a lower mean trophic level [@Lopez-Lopez2022] (Table 1).

In the Antarctic realm, Potter Cove is a ~9 km2 fjord with a depth range of 0 - 200 m located at 25 de Mayo/King George Island (62°14’S, 58°38’W, South Shetland Islands) on the West Antarctic Peninsula (Figure 1). The cove, bordered by the Fourcade Glacier, is divided into three areas: a) the internal cove, a high glacier-influenced, soft sediment zone with a 50 m maximum depth; b) the central cove, a mixed substrate area with low meltwater influence and an 80 m maximum depth; and c) the external cove, ice-free for 60 years with a 185 m maximum depth and rocky substrate [@Jerosch2018]. Potter Cove's high-latitude location results in variable environmental conditions due to photoperiod length seasonality. Sea ice often covers this area in winter [@Schloss2012]. With low phytoplankton biomass, macroalgae, and microphytobenthos are likely the primary food sources for secondary benthic production [@Quartino2008].
The Potter Cove food web includes 110 nodes and 649 interactions, with an asymmetric degree distribution, and a connectance value of 0.05 (Table 1). It presents a modular structure (groups of species interact more strongly with each other than with species belonging to other groups), that’s positively associated with stability, since perturbation can be retained within modules [@Rodriguez2022].

Located between 74 and 78ºS, the high Antarctic Weddell Sea shelf spans approximately 450 km from East to West [@Jacob2011] (Figure 1). The water depth ranges from 200 to 500 m, with shallower regions being covered by continental ice that forms the coastline along the eastern and southern parts of the Weddell Sea. Within this shelf area, exists a complex three-dimensional benthic habitat characterized by substantial benthic biomasses and an intermediate to high diversity when compared to benthic boreal communities [@Teixido2002].
The Weddell Sea food web exhibits a high level of network complexity, featuring the greatest number of nodes (490) and trophic interactions (16041) among the analyzed food webs in this review (Table 1). Its connectance (0.07) and other structural properties are intermediate compared to the other food webs. Recently, the interaction strengths of this food web were estimated, revealing the presence of numerous weak and few strong interactions, which is consistent with findings in other complex food webs [@Nilsson2016]. Notably, this asymmetric distribution of interaction strength is likely to promote community persistence.

>>>>

# The response of trophic interaction networks to multiple stressors along a large-scale latitudinal range in the Southern Hemisphere
## Running Title: Marine stressors in the Southern Hemisphere

@@

Ecological networks offer valuable insights into community structure, key species identification, and ecosystem management. Understanding how these networks respond to global change stressors is of increasing interest, especially along geographical gradients.
This review summarizes potential stressor responses in marine food webs from the Southwest Atlantic to the Antarctic (45 - 78ºS), encompassing areas such as San Jorge Gulf, Beagle Channel, Burdwood Bank, Scotia Sea, Potter Cove, and the Weddell Sea in Antarctica. The objectives are: 1) to describe the structure of marine food webs along this latitudinal axis using a network approach; 2) to identify predominant global change-related stressors affecting each ecosystem; and 3) to summarize observed food web changes and hypothesize on stressor impacts. The effects of stressors were primarily reviewed at the species level.
Alternative hypotheses for each study area were formulated considering: a) main stressors; b) impacted parameters; c) node-level species properties; and d) network-level food web properties. Global warming emerges as the most common stressor among the studied areas across the latitudinal gradient, except in the Beagle Channel and Burdwood Bank, where alien species introduction and fisheries are more influential. We offer a series of alternative hypotheses on how warming may affect the food webs. This review emphasizes the benefits of using a network approach to understand and predict stressor effects in Southern Hemisphere marine ecosystems. This approach provides a holistic understanding of ecosystems, which enhances our ability to identify key species and their interactions, offering insights for ecosystem management and conservation in the face of global change stressors.

Keywords: stressors, global warming, food webs, latitudinal gradient, Southern Hemisphere

@@

The application of a network perspective has emerged as a powerful tool to tackle the complexity of species interactions, facilitating a better understanding of the structure and functioning of ecosystems [@Belgrano2005; @Thompson2012]. Trophic networks (or food webs) allow identifying properties and key species that may be crucial for ecosystem stability, and hence important for ecosystem management and biodiversity conservation [@Thompson2012].
There is a growing interest in understanding how ecological networks respond to environmental and anthropogenic stressors along geographical gradients [@Bauer2022; @Cirtwill2015]. Yet, only a few studies have described variation in food web structure along latitudinal gradients in marine ecosystems. The few that have come from the Global North [@Wood2015; @Kortsch2019; @Pecuchet2020], whereas no studies, nor meta-analyses, on geographical variation in marine food webs exist for the Global South (Southern Hemisphere).

Here we review for the first time stressor responses of marine food webs for areas located along a Southwest Atlantic to Antarctic gradient (45 - 78ºS, Figure 1). We considered proven and expected changes in food webs driven by global change-related stressors in selected areas along this large-scale latitudinal range.
The aim of this review is threefold: 1) describe the complexity and structure of marine food webs along the Southwest Atlantic to Antarctic axis from a network perspective; 2) identify the ongoing global change-related stressors for each marine ecosystem containing the food webs; and 3) review proven food web changes and elaborate on a series of alternative hypotheses on how the identified stressors might affect food web features (energy flow, stability and topology), combining information on node- and network-level properties. 
To identify the main stressors, we performed an exploratory and configurative review instead of a systematic review [@Haddaway2015], because our objective was to formulate evidenced-based hypotheses rather than providing an exhaustive list of species-level stressors for each ecosystem.
Finally, we suggest which additional data and analyses are needed to gain insights into the stressors’ effects on food web properties in the Southwest Atlantic to Antarctic region.

________________

## 2. Assessment of the food web structure and stressors in the Southwest Atlantic - Antarctic region

The Southwest and the Atlantic sectors of the Southern Ocean comprise one of the most biologically productive regions of the world's oceans [@Acha2004; @Latorre2023]. The referred region extends from San Jorge Gulf (45ºS) in the Patagonian shelf to the Weddell Sea (78ºS) in the Southern Ocean and covers a well-connected oceanic area [@Matano2010; @Guihou2020].

Throughout this region, many studies address the ecology of predator-prey relationships but most of these studies are restricted to a few interacting species or certain components of the food web such as fish, crustaceans, and benthos (e.g., @Vinuesa2007, @Padovani2012, @Pasotti2015, @Riccialdelli2020, @Saunders2019, @Pasti2021, @Cordone2022). Only few studies consider the complexity of trophic interactions within highly resolved food webs containing hundreds of species, ranging from phytoplankton, zooplankton, benthos, fish, seabirds to whales (but see @Jacob2011, @Marina2018, @Funes2022, @Lopez-Lopez2022, @Rodriguez2022, @Marinainrev). Neglecting this complexity might lead to a misunderstanding of the structure and functioning of the ecosystems and ultimately reduce the ability to predict ecosystem responses to perturbations [@Montoya2009].

The Southwest Atlantic - Antarctic marine biota has been and is currently subjected to a variety of stressors (e.g., sea warming, glacial retreat, ice changes, acidification, species invasion, fisheries, and contamination). Here, we consider a stressor as a global change variable if it causes a quantifiable change, irrespective of its direction (increase or decrease), in a biological response [@Orr2020]. This aligns with the framework established during the early years of the International Geosphere-Biosphere Programme (IGBP) for classifying driving forces. These driving forces are grouped into those that modify the composition of the Earth's fluid envelopes (e.g., sea warming, acidification), and those that originate as local processes but attain global importance due to their widespread impact (e.g., species invasion, fisheries) [@Vitousek1992; @Sage2020].
Although it is well-known that multiple stressors act in concert at any given time (e.g., warming and fishery; acidification and contaminants) [@Gutt2021], to date stressor assessments have been performed individually. Moreover, the potential for interactive effects of two or more stressors (synergy or antagonism) [@Cote2016] is almost unknown for the region [@Rowlands2021].

In this review, we included six large marine areas in the Southern Hemisphere for which highly-resolved empirical food webs in terms of species diversity and trophic interactions have been previously described: (1) San Jorge Gulf (~45 - 47ºS, 65 - 68ºW) [@Funes2022], (2) Beagle Channel (~54ºS, 68ºW) [@Rodriguez2022], (3) Burdwood Bank (~54ºS, 59ºW) [@Marinainrev], (4) Scotia Sea (~48 - 58ºS, 50ºW) [@Lopez-Lopez2022], (5) Potter Cove (62ºS, 58ºW, Antarctica) [@Marina2018], and (6) Weddell Sea (74 - 78ºS, 30ºW, Antarctica) [@Jacob2011; @Marina2024] (Figure 1). The geographic locations of these marine biomes span from temperate Atlantic waters to Antarctic, and are exposed to global change stressors (e.g., sea warming, glacial retreat, fishery, pollution) [@Sage2020], of which some are locally more important, e.g., fisheries in the San Jorge Gulf or glacial retreat in the Antarctic.

In the following subsections, we describe the complexity and structure of the mentioned marine food webs, identify the main stressors, the species (or trophic species), and the parameters affected. Table 1 summarizes the properties of the marine food webs (number of nodes, trophic interactions, connectance, path length, mean trophic level, and omnivory); Table 2 provides a list of the main stressors affecting species inhabiting each area, considering: stressors’ nature, species and parameter affected, and locality.

@@

### 2.1 San Jorge Gulf

San Jorge Gulf is the northernmost study area considered in this review (Figure 1), located from Cabo dos Bahias (44°55'S) to Cabo tres Puntas (47°06'S). It is a partially enclosed basin spanning approximately 34,000 km^2 and ~100 m of maximum depths. The gulf’s productivity supports large invertebrate and vertebrate fisheries [@Gongora2012], as well as marine mammal and seabird populations [@Yorio2009]. Two prominent frontal systems meet in the shallower northern and southern ends of the gulf (with depths ~40 m), which are also the places of highest productivity [@Glembocki2015].

The San Jorge Gulf food web contains 165 nodes (trophospecies) and 1015 trophic interactions with a connectance of 0.04. The percentage of top predators is 16%, 78% of the nodes are intermediate and 6% are basal; 60% of the consumers are omnivorous (Table 1). The most connected nodes are the Argentine red shrimp *Pleoticus muelleri*, the squat lobster *Gimothea gregaria*, squids (*Illex argentinus* as dominant species), and Amphipoda. Notably, these nodes are positioned at mid-trophic levels in the food web (3, 2.5, 3.6, and 2, respectively) [@Funes2022].

The gulf is subject to several global change stressors (Table 2), especially fisheries [@Gonzalez-Zevallos2006; @Gonzalez-Zevallos2011; @Galvan2022]. By discarding species, trawl fisheries add new trophic interactions to the food web which results in decreasing trait variability and stability of the system [@Funes2022; @Rincon-Diaz2021]. Moreover, this has changed the availability of resources to some predators. For example, the Argentine hake *Merluccius hubbsi*, one of the main bycatch species, has become a prey item of non-diving seabirds, like the kelp gull *Larus dominicanus* [@Gonzalez-Zevallos2006] and reef fish [@Funes2019]. Although juveniles of *M. hubbsi* is largely the main bycatch species, 29 cartilaginous and 69 bony local fish species were also registered as incidental catch between 2005 and 2014 [@Bovcon2013; @RuibalNunez2020]. This level of impact has triggered a shift in the functional diversity of fish assemblages in the gulf, homogenizing their trophic function [@Rincon-Diaz2021], for example in the sites with the historical highest fishing activity, the fish assemblage experienced a relative increase in midwater fusiform species and lost variability in their depth range [@Rincon-Diaz2021]. Other functional changes were a decrease in the maximum sizes of individual fish, together with a drop in elasmobranch biomass and an increase in crustacean biomass [@Funes2020]. The significant increase in crustacean biomass was mainly due to the increase in the Argentine red shrimp and the squat lobster [@Funes2020]. These species have become the most important prey for the most abundant fish in the area, the Argentine hake and the narrownose smooth-hound *Mustelus schmitti* [@Belleggia2017; @Pasti2021]. However, since commercial fisheries ceased their activities in the San Jorge Gulf in 2015 (Annex I, Resolution CFP No 7/2018), the effects mentioned above from trawl fisheries on the structure and function of the ecosystem may have changed since then. Nowadays only small-scale trawling from artisanal fishery is practiced in the gulf.
Sea warming is another important stressor in the San Jorge Gulf, because of the southward shifts of northern fish populations to the area [@Galvan2022]. Signs of sea warming have been detected in the gulf and nearby regions. Anomalous bottom temperature increments of up to 2°C have been observed in the gulf since 2000 [@Isla2024], and a positive trend in sea surface temperature of 0.3°C/0.4°C per decade over a 30-year series [@Saraceno2022; @Risaro2022]. Additionally, a positive trend in sea surface temperature (~0.5 °C increase in the last 20 years) has been reported for the Nuevo and San Jose Gulfs, both located close to and northward of the San Jorge Gulf [@Williams2023]. The region is especially prone to be affected by climate-driven shifts in species ranges because it is situated in the ecotone of two biogeographic provinces, the Argentine (30°S - 44°S) and the Magellanic (43°S - 55°S) [@Balech2008]. In addition to tropicalization from northern fish populations, alien species are also documented to affect demersal assemblages of fish and macroinvertebrates in the gulf [@Galvan2022].
Finally, the gulf is exposed to urban and industrial pollution [@Verga2020]. In particular, there is an oil monobuoy from which oil manipulation and general oil transport along the Patagonian coast registered several oil spills and chronic oil discharges [@Garcia-Borboroglu2008]. Other marine systems impacted by the oil spill have shown extraordinary mortality events for cetacean [@Litz2014], sea turtles [@Wallace2017], and important decrease in marine bird populations [@Irons2000] sea otters [@Garrott1993] and macroalgae [@Southward1978], with cascading effects throughout the food web [@Peterson2001].

### 2.2 Beagle Channel

Located at a more southern latitude (~54ºS, 68ºW), the Beagle Channel (Figure 1) is an interoceanic passage at the southernmost tip of South America, spanning 240 km in length and 5 km in width (~1200 km^2), with a depth range of 0 - 140 m. It features complex coastlines, varying bathymetry, a prevailing west-to-east circulation pattern, and a significant longitudinal gradient of glacial freshwater discharge [@Schloss2023]. @Bruno2023 suggests that locally produced suspended particulate organic matter (mostly composed of phytoplankton) and organic matter accumulated in the sediments (macroalgae-originated detritus) are the primary food sources for the marine ecosystem, as opposed to allochthonous materials.

The Beagle Channel food web includes 145 nodes and 1115 trophic interactions with a connectance of 0.05 (Table 1). This food web is hypothesized to have a wasp-waist structure (i.e., the structure and dynamics of the ecosystem are regulated primarily by mid-trophic level species), where the following species play a crucial role in the dynamics of the ecosystem: Fueguian sprat *Sprattus fuegensis*, longtail southern cod *Patagonotothen ramsayi*, black southern cod *P. tessellata*, frogmouth *Cottoperca trigloides*, and squat lobster *G. gregaria* [@Riccialdelli2020]. Moreover, the squat lobster has been identified as a species linking food web modules and connecting the entire food web [@Rodriguez2022].

A main local stressor in the Beagle Channel is the introduction of invasive species (Table 2). Salmonidae were introduced to Tierra del Fuego in the 1930s. Especially, the Chinook salmon *Oncorhynchus tshawytscha* causes concern. The detection of this species in Tierra del Fuego dates back to April 2006, and its population has been expanding since then [@Nardi2019]. Being a top predator, the Chinook salmon can compete for resources with several native predators in the Beagle Channel [@Correa2008], but also prey on them. This is the case for Notothenioids (*Eleginops maclovinus*, *Patagonotothen tessellata*, *P. cornucola*, *P. sima*, *Paranotothenia magellanica*, *Harpagifer bispinis*), the Atherinidae *Odontesthes smitti* and *O. nigricans*, the Fuegian sprat, and larvae of king crabs (*Lithodes antarcticus* and *Paralomis granulosa*) [@Fernandez2010]. @Ciancio2008 observed that Chinook salmon primarily feeds on sprats in the Southern Patagonian Shelf Ecosystem area and display trophic levels comparable to those of intermediate-sized fish and cephalopod predator species, showing significant dietary overlap with the Magellanic penguin *Spheniscus magellanicus*. Another potential competitor for Chinook salmon in the Beagle Channel, sharing a similar diet, is the native Commerson's dolphin *Cephalorhynchus commersonii* [@Riccialdelli2013].
In the Beagle Channel, other stressors include contaminants like metals, perfluorinated compounds, hydrocarbons, and microplastics found in animal tissue (algae, mussel, fish, and guano) and sediments [@Duarte2011; @Llorca2012]. Carbon and carbohydrate levels in Ushuaia Bay's (54°48′S, 68°18′W) surface sediments are similar to hypertrophic ecosystems (i.e., greater input of nutrients than an eutrophic situation), along with hydrocarbons and heavy metals linked to port and industrial activities [@Gil2011; @Commendatore2012]. @Ferreira2021 showed that the black southern cod in Ushuaia Bay may be exposed to endocrine-disrupting compounds from urban and industrial pollution.
Moreover, @Perez2020 and @Ojeda2021 found microplastics in the Chilean mussel *Mytilus chilensis* and the sea snail *Nacella magellanica*, respectively. These studies link pollutants to port and industrial activities of Ushuaia; we therefore expect the area of the Beagle Channel closest to the city to be the most affected by contaminants. 
Contaminants potentially spread through the food web, magnifying the risk to higher trophic level taxa. In the Beagle Channel and Atlantic coast of Tierra del Fuego food webs mercury concentrations rose with benthivory, and higher mercury was found in phytoplankton and the squat lobster. This latter species connects pelagic and benthic habitats, so any disruption to them could disrupt the food web [@Rodriguez2022]. @Dodino2022 found the highest mercury levels in Magellanic penguin feathers from offshore colonies in Tierra del Fuego. Recently, Ushuaia's kelp forests have also seen changes in composition along with a decrease in biodiversity as a result of urban pollution [@Kaminsky2024]. It is important to note that sea warming might be a less significant stressor in the Beagle Channel compared to the other areas, as global warming is melting glaciers and the influx of cold freshwater is actually slightly cooling the seawater [@Saraceno2022].

### 2.3 Burdwood Bank

The so-called Burdwood Bank ecosystem comprises Marine Protected Areas Namuncurá - Burdwood Bank I and II (~54ºS, 59ºW), meaning the shallow submarine plateau named Burdwood Bank with a 200 m isobath boundary, and a deep slope that reaches 4000 m in depth, respectively [@APN2022] (Figure 1). Physical features in the plateau are fairly stable, with salinity averaging 34 all year round and temperature ranging between 4 and 8ºC [@Acha2004]. The plateau is surrounded by steep flanks of up to 4000 m depth, protected by the Marine Protected Area Namuncurá - Burdwood Bank II (32,000 km^2). Intense upwelling and mixing occur in relation to the slope, entraining deep nutrient-rich waters into the photic layer [@Matano2019], and resulting in a fairly homogeneous water column both spatially and temporally [@Matano2019].

The Burdwood Bank food web comprises 379 nodes and 1788 interactions, with a connectance of 0.01, and an asymmetric degree distribution (i.e. most of the species have a relatively low number of interactions and few species concentrate most of them). Almost half of the consumers are omnivores (48%), and the network displays a small-world pattern [@Marinainrev] (Table 1). This network pattern implies a rapid spread of a perturbation (e.g., contaminant, population fluctuation, local extinction) throughout the food web due to a short distance among species but, at the same time, a greater resistance caused by the confinement of perturbations mainly within subnetworks as a result of a high clustering coefficient [@Bornatowski2017; @Dormann2017].

The main stressors reported for the ecosystem of the Marine Protected Areas Namuncurá within Burdwood Bank I and II are related to human activities such as fisheries and contamination (Table 2). Several fisheries targeting demersal fish operate in the vicinity and within the Marine Protected Areas. The fishery of the Patagonian toothfish *Dissostichus eleginoides* has gained prominence in recent years [@Allega2020; @Gorini2021]. Although these are regulated by the Argentinean government, incidental bycatches do occur, where demersal fish of the genera *Coelorinchus* and *Macrourus*, seabirds and benthic invertebrates (30+ taxa) are the most common bycatches [@Gaitan2016; @Martinez2022]. Noteworthy, among the invertebrates caught, eight species are indicator taxa of vulnerable marine ecosystems [@Gaitan2016; @Schejter2021]. However, independent assessments of these bycatches suggest no significant impact on the populations [@Gaitan2016; @Martinez2022]. Besides bycatches, seabirds are being affected by fisheries due to discards, altering their diet; the most frequently encountered species are black-browed albatross *Thalassarche melanophris*, Southern giant petrel *Macronectes giganteus*, Cape petrel *Daption capense*, Southern royal albatross *Diomedea epomophora*, Northern giant petrel *M. halli*, and white-chinned petrel *Procellaria aequinoctialis* [@Tamini2023]. Nevertheless, there is a lack of knowledge considering their role in the ecosystem and the potential joint effect of both target fish and bycatch in a broader food web framework.
Contaminants such as microplastics and mercury occur in the water column of the Burdwood Bank ecosystem [@Cossi2021; @DiMauro2022; @Fioramonti2022]. Microplastics are distributed along the entire water column, from surface to deep waters (3-2450 m) [@DiMauro2022]. More importantly, microplastics were found in soft tissues of benthic macroinvertebrates (sea stars *Henricia obesa* and *Odontaster penicillatus*) and benthopelagic fish (*Patagonotothen guntheri* and *P. ramsayi*), which not only incorporated the contaminant from the environment through their filter-feeding system but could also get it indirectly from prey organisms already containing plastics in their tissues [@Cossi2021]. Notably, one of the contaminated species is the longtail southern cod *P. ramsayi*. This species is part of the core group of species that drive the ecosystem through the suggested wasp-waist control [@Riccialdelli2020].
Mercury transfer and biomagnification are occurring at a greater pace than near coastal areas such as the Beagle Channel [@Fioramonti2022]. The Fuegian sprat *S. fuegensis*, a pelagic fish with a mid-trophic level in the food web, has the highest levels of mercury recorded [@Fioramonti2022]. Considering the wasp-waist control of the Fuegian sprat in the food web [@Riccialdelli2020], a rapid and widespread contamination to the top predators is expected [@Fioramonti2022].
Despite evidence of warming of surface, mid-water and bottom layers (100 m) in Burdwood Bank [@Franco2020], studies on the oceanographic aspects of the system are lacking. Warming impacts on the species and trophic interactions in Burdwood Bank are also unknown.

### 2.4 Scotia Sea

The Scotia Sea is a deep-sea basin (~48 - 58ºS, 50ºW), delimited by the Drake Passage to the West and by the island complex of the Scotia Arc to the North, East, and South, with an approximate extension of 1.5 x 106 km^2 and a depth range of 0 - 3000 m [@Murphy2006] (Figure 1). Its oceanography is dominated by the Antarctic Circumpolar Current, which is spatially structured by frontal systems [@Whitworth1980]. The South Antarctic Circumpolar Current Front subdivides the Scotia Sea into two biogeographic regions: the Northern Scotia Sea, characterized by higher and more variable temperatures, and the Southern Scotia Sea, characterized by lower and more stable temperatures and influenced by seasonal sea ice [@Raymond2011].

Analysis of the Northern and Southern Scotia Sea food webs shows that the former is relatively more complex in terms of number of species and links than the latter: with more species richness (218 vs 192) and interactions (10008 vs 7241) and a slightly higher connectance (0.21 vs 0.20). Mean path length is shorter in the Northern Scotia Sea food web, whereas the Southern Scotia Sea network displays a greater proportion of omnivores and a lower mean trophic level [@Lopez-Lopez2022] (Table 1).

The Scotia Sea is a vast and heterogenous oceanic region, where especially the zone around South Georgia island represent an area of interest, here referred to as ‘Northern Scotia Sea’. The majority of studies analyzing the stressors’ effects come from this area (e.g., @Murphy2007; @Whitehouse2008; @Trathan2023) (Table 2).
The main stressor of the Scotia Sea is commercial fisheries. Krill fishery on *Euphausia superba* not only impacts the targeted species, but also the many predators that depend on it as a food source [@Hilborn2017]. Yet, data currently available from monitoring of krill and its predators remain insufficient, hence identifying the potential fishery impacts on the ecosystem is difficult [@Trathan2021]. Apart from krill fishery, two other commercial fisheries operate in the Scotia Sea, targeting Patagonian toothfish species *D. eleginoides* and *Dissostichus mawsoni*. The *D. eleginoides* stock is linked to the stock at South Georgia (‘Northern Scotia Sea’) [@Collins2010], while the *D. mawsoni* stock is linked to the Antarctic continental shelf (‘Southern Scotia Sea’) [@Soeffker2022]. Recently, @Trathan2023 has identified several concerns regarding aspects of these fisheries, highlighting that it is crucial to recognize ongoing changes such as the recovery of baleen whales and climate change. These factors increase uncertainty for fisheries managers, necessitating direct consideration in the management of harvested and dependent species.

The Scotia Sea has experienced one of the largest levels of warming within the polar regions [@Whitehouse2008]. Together with the Southern Annular Mode anomalies, this has caused a long-term decrease in krill abundance; more pronounced in the northern than in the southern Scotia Sea [@Murphy2007]. Over the past 90 years, the Antarctic krill *E. superba* also showed an increase in mean body length [@Atkinson2019], likely altering predator-prey interactions. Moreover, it allows krill to reach colder feeding grounds near the seabed, with the potential to link krill to unexpected predators such as demersal fish [@Schmidt2011]. Because of warming, the krill distribution has shifted southward [@Atkinson2019], changing the food web from a krill-based to a non-krill-based, where myctophid fish and squid are playing important roles as key species in a wasp-waist controlled system [@Saunders2019]. While the krill distribution has shifted southward, the most abundant calanoid copepods have maintained their distribution [@Tarling2018].
Mercury transfer and biomagnification are current processes occurring in the Scotia Sea, where the total concentration of contaminants increase with trophic level and are highest in notothenioid and myctophid fish (e.g. *D. eleginoides*, *Gymnoscopelus nicholsi*), and seabirds [@Seco2021]. During years of low Antarctic krill abundance, predators must deal with both the stress of reduced prey availability and the concurrent rise in mercury exposure [@Seco2021].

### 2.5 Potter Cove (Antarctica)
